---
title: "Multiple regression"
format: html
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)   # For ggplot, dplyr, and friends
library(moderndive)  # For get_regression_table(), geom_parallel_slopes(), etc

# Load the World Happiness Report data and only keep 2024
world_happiness <- read_csv("data/world_happiness.csv") |>
  filter(year == 2024) |>
  mutate(
    # GDP per capita in thousands of dollars, so that a "one unit increase" is
    # $1,000 instead of $1
    gdp_per_cap_1000 = gdp_per_cap / 1000,
    # Put the income groups in order (otherwise R puts them in alphabetical order)
    income_group = factor(
      income_group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    )
  )
```

## World happiness

This data comes from the [World Happiness Report](https://www.worldhappiness.report/) (happiness scores) and the [World Bank](https://data.worldbank.org/) (everything else). Each row is a country in 2024.

- `country`: Country name
- `year`: Year
- `continent`: Africa, Americas, Asia, Europe, or Oceania
- `region`: World Bank region (East Asia & Pacific, Sub-Saharan Africa, etc.)
- `income_group`: World Bank income group (Low income, Lower middle income, Upper middle income, High income)
- `iso3c`: Three-letter country code
- `happiness`: Average life evaluation score, on a scale of 0–10 (higher = happier)
- `life_expectancy`: Average life expectancy at birth, in years
- `gdp_per_cap`: GDP per capita (constant 2015 US dollars)
- `gdp_per_cap_1000`: GDP per capita, in thousands of dollars
- `internet_use_prop`: Percent of the population that uses the internet (0–100)
- `urban_prop`: Percent of the population that lives in cities (0–100)
- `women_parliament_prop`: Percent of seats in the national parliament held by women (0–100)
- `unemployment_rate`: Percent of the labor force that is unemployed (0–100)
- `population_millions`: Population, in millions

```{r}
world_happiness
```


## Our turn #3: Happiness, wealth, internet access, and continents

Is national happiness explained by wealth, internet access, *and* continent?

Fit a model with `lm()` that explains `happiness` with `gdp_per_cap_1000`, `internet_use_prop`, and `continent`, and look at the coefficients with `get_regression_table()`:

```{r}
# model_happiness <- lm(...)
```

Write out the model as an equation (fill in the numbers):

$$
\widehat{\text{happiness}} = b_0 + b_1 \times \text{GDP}_{1000} + b_2 \times \text{internet} + b_3 \times \text{Americas} + b_4 \times \text{Asia} + b_5 \times \text{Europe} + b_6 \times \text{Oceania}
$$

Interpret the intercept:

> In an African country with a GDP per capita of \$0 and 0% internet use, ...

Interpret the coefficient for GDP per capita:

> Taking internet use and continent into account, a \$1,000 increase in GDP per capita is associated with a ___ ...

Interpret the coefficient for internet use:

> ...

Interpret the coefficient for Europe:

> ...

Bonus: plot `happiness` and `gdp_per_cap_1000` with one line for each continent using `geom_parallel_slopes(se = FALSE)` (this plot ignores internet use, so the lines won't match the model exactly!):

```{r}
ggplot(data = world_happiness)
```


## Your turn #3: Happiness, health, jobs, and income

Is national happiness explained by health, jobs, and income?

Fit a model that explains `happiness` with `life_expectancy`, `unemployment_rate`, and `income_group`, and look at the coefficients with `get_regression_table()`:

```{r}
# model_income <- lm(...)
```

What's the base case for `income_group`?

> ...

Interpret the coefficient for life expectancy:

> ...

Interpret the coefficient for unemployment:

> ...

Interpret the coefficient for high income:

> ...


## Playground

Build your own model! Pick an outcome and a few explanatory variables (try mixing continuous ones like `urban_prop`, `women_parliament_prop`, or `population_millions` with categorical ones like `region` or `income_group`). Fit the model, look at the coefficients, and interpret each one.

```{r}
# model_playground <- lm(...)
```

Interpretations:

> ...
