---
title: "Simple linear regression"
format: html
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)   # For ggplot, dplyr, and friends
library(moderndive)  # For get_correlation(), get_regression_table(), etc

# Load the World Happiness Report data and only keep 2024
world_happiness <- read_csv("data/world_happiness.csv") |>
  filter(year == 2024)
```

## World happiness

This data comes from the [World Happiness Report](https://www.worldhappiness.report/) (happiness scores) and the [World Bank](https://data.worldbank.org/) (everything else). Each row is a country in 2024.

- `country`: Country name
- `year`: Year
- `continent`: Africa, Americas, Asia, Europe, or Oceania
- `region`: World Bank region (East Asia & Pacific, Sub-Saharan Africa, etc.)
- `income_group`: World Bank income group (Low income, Lower middle income, Upper middle income, High income)
- `iso3c`: Three-letter country code
- `happiness`: Average life evaluation score, on a scale of 0–10 (higher = happier)
- `life_expectancy`: Average life expectancy at birth, in years
- `gdp_per_cap`: GDP per capita (constant 2015 US dollars)
- `internet_use_prop`: Percent of the population that uses the internet (0–100)
- `urban_prop`: Percent of the population that lives in cities (0–100)
- `women_parliament_prop`: Percent of seats in the national parliament held by women (0–100)
- `unemployment_rate`: Percent of the labor force that is unemployed (0–100)
- `population_millions`: Population, in millions

```{r}
world_happiness
```


## Our turn #1: Happiness and life expectancy

Does a country's life expectancy predict its happiness?

Make a scatterplot with `life_expectancy` on the x-axis and `happiness` on the y-axis:

```{r}
ggplot(data = world_happiness)
```

Find the correlation between the two with `get_correlation()`:

```{r}
world_happiness
```

Fit a model with `lm()` and look at the coefficients with `get_regression_table()`:

```{r}
# model_life_exp <- lm(...)
```

Interpret the slope:

> On average, a one-year increase in life expectancy is associated with a ___ ...

Interpret the intercept:

> In a country where life expectancy is 0 years, ...


## Your turn #1: Happiness and internet access

Does access to the internet predict national happiness?

Make a scatterplot with `internet_use_prop` on the x-axis and `happiness` on the y-axis:

```{r}
ggplot(data = world_happiness)
```

Find the correlation between the two with `get_correlation()`:

```{r}
world_happiness
```

Fit a model with `lm()` and look at the coefficients with `get_regression_table()`:

```{r}
# model_internet <- lm(...)
```

Interpret the slope:

> On average, a one percentage point increase in internet use is associated with a ___ ...

Interpret the intercept:

> ...


## Our turn #2: Happiness and continents

Is happiness different across continents?

Make a plot that shows `happiness` across each `continent` (try `geom_point()` with `position = position_jitter(width = 0.15)`, or `geom_boxplot()`):

```{r}
ggplot(data = world_happiness)
```

Find the average happiness in each continent with `group_by()` and `summarize()`:

```{r}
world_happiness
```

Fit a model with `lm()` and look at the coefficients with `get_regression_table()`:

```{r}
# model_continent <- lm(...)
```

What's the base case?

> ...

Interpret the intercept:

> ...

Interpret the coefficient for Europe:

> On average, happiness in Europe is ___ ...


## Your turn #2: Internet access and continents

Is internet access different across continents?

Make a plot that shows `internet_use_prop` across each `continent`:

```{r}
ggplot(data = world_happiness)
```

Find the average internet use in each continent with `group_by()` and `summarize()`:

```{r}
world_happiness
```

Fit a model with `lm()` and look at the coefficients with `get_regression_table()`:

```{r}
# model_internet_continent <- lm(...)
```

Interpret the intercept:

> ...

Interpret the coefficient for Asia:

> ...
