---
title: "Getting started with R and Positron"
format: 
  html: default
  typst: default
  docx: default
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(gapminder)

gapminder_2007 <- filter(gapminder, year == 2007)
```

## Quarto

This is a [Quarto](https://quarto.org/) file (it has a .qmd file extension). When you execute code within the file, the results appear beneath the code. 

R code goes in **code cells** or **code chunks**, denoted by three backticks. Try executing this cell by clicking the *Run* button (a small green triangle) within the cell or by placing your cursor inside it and pressing *Ctrl+Shift+Enter* (or *Cmd+Shift+Enter* on Mac). 

```{r}
ggplot(data = gapminder_2007) +
  geom_point(mapping = aes(x = gdpPercap, y = lifeExp, color = continent))
```

## Add a new code cell

Add a new code cell by clicking the *Insert Code Cell* button on the toolbar or by pressing *Cmd/Ctrl+Option+I*. 

Put 2 + 2 in your new code cell and run it.

PUT A NEW CELL HERE


## Rendering Quarto files

We'll use Quarto files as notebooks as we learn because we can record text, code and output. 

Quarto files are also a publication format. Try hitting the "Preview" button in the toolbar above. R runs all the code in the document from top to bottom, it collects the output and puts the code, text and output together in an HTML document---you should see it as `01-getting-started.html` in the Files pane. This document is a great way to record or share your work (you can also render to PDF or Word documents).

## Assigning variables

What's the difference between the code in this cell:

```{r}
filter(gapminder, lifeExp > 70)
```

And the code in this cell?

```{r}
high_life_exp <- filter(gapminder, lifeExp > 70)
```

## Functions

Look at the help page for `seq`

Add a cell here that uses `seq()` to create a list of numbers from 5 to 100, spaced by 5 (so 5, 10, 15, 20, etc.)

CELL HERE


## Syntax gone wrong

```{r}
#| error: true
sd(pull(.data = starwars, var = mass)
```

```{r}
#| error: true
my_name <- "Andrew'
```

```{r}
#| error: true
pull(.data = "starwars", var = height)
```
