---
title: "Exploring charitable activity in the Middle East"
format: 
  html: default
  typst: default
  docx: default
---

## Background

Wave 8 of the Arab Barometer (2023-2024) was conducted in seven different countries: Iraq, Jordan, Kuwait, Lebanon, Mauritania, Morocco, Palestine, and Tunisia. See [their questionnaire code book here](https://www.arabbarometer.org/survey-data/data-downloads/) for complete details of every question they asked.

## Donations to charity

Question 266 asked:

> In a typical month, do you donate money to a charity or those in need?

I want to know if there are country-level differences in charitable activity.

## Data loading and cleaning

First, I'll load the data and clean it up a bit:

```{r}
#| label: load-data
#| warning: false
#| message: false

library(tidyverse)
library(haven)

barometer <- read_stata("data/ArabBarometer_WaveVIII_English_v3.dta") |> 
  mutate(
    COUNTRY = as_factor(COUNTRY), 
    Q266 = as_factor(Q266)
  )
```

## Proportion of people donating by country

Here's the proportion of people who responded yes in each country:

```{r}
donation_by_country <- barometer |>
  group_by(COUNTRY) |>
  summarize(
    total_responses = n(),
    yes_count = sum(Q266 == "Yes", na.rm = TRUE),
    prop_yes = mean(Q266 == "Yes", na.rm = TRUE)
  )
donation_by_country
```

Let's look at a picture of that:

```{r}
donation_by_country |>
  mutate(COUNTRY = fct_reorder(COUNTRY, prop_yes)) |>
  ggplot(aes(x = prop_yes, y = COUNTRY)) +
  geom_col() +
  scale_x_continuous(labels = scales::label_percent()) +
  labs(x = "Proportion who donated to charity last month", y = "Country")
```
