---
title: "Visualizing data with {ggplot2}"
format: 
  html: default
  typst: default
  docx: default
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)  # For ggplot2, dplyr, and friends

# Load data and remove rows where the "sex" column is missing
penguins <- read_csv("data/penguins.csv") |> 
  drop_na(sex)
```

## Figuring out what's in the data

```{r}
penguins
```

## Your turn #1

Run the code from the the slide to make a graph.

```{r}
ggplot(data = penguins) + 
  geom_point(mapping = aes(x = bill_len, y = body_mass))
```

## Your turn #2: ggplot playground

Add `color`, `size`, `alpha`, or `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = penguins) + 
  geom_point(mapping = aes(x = bill_len, y = body_mass, color = "blue"))
```

## Your turn #3: Switching geoms

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(data = penguins) +
  geom_boxplot(mapping = aes(x = species, y = body_mass))
```

## Your turn #4: Histograms

Make a histogram of the `body_mass` variable from `penguins`.

```{r}
ggplot(data = penguins)
```

## Your turn #5: Density plots

Make a density plot of `body_mass` filled by `species` with 75% transparency.

```{r}
ggplot(data = penguins) 
```

## Your turn #6: Complex plots

Predict what this code will do. Then run it.

```{r}
ggplot(data = penguins) + 
  geom_point(aes(x = bill_len, y = body_mass)) +
  geom_smooth(aes(x = bill_len, y = body_mass))
```

## Takeaways

You can use this code template to any kind of graph with {ggplot2}.

```{.r}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```
