---
title: "Multiple regression"
format: html
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)   # For ggplot, dplyr, and friends
library(moderndive)  # For get_regression_table(), geom_parallel_slopes(), etc

# Load the World Happiness Report data and only keep 2024
world_happiness <- read_csv("data/world_happiness.csv") |>
  filter(year == 2024) |>
  mutate(
    # GDP per capita in thousands of dollars, so that a "one unit increase" is
    # $1,000 instead of $1
    gdp_per_cap_1000 = gdp_per_cap / 1000,
    # Put the income groups in order (otherwise R puts them in alphabetical order)
    income_group = factor(
      income_group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    )
  )
```

## World happiness

This data comes from the [World Happiness Report](https://www.worldhappiness.report/) (happiness scores) and the [World Bank](https://data.worldbank.org/) (everything else). Each row is a country in 2024.

- `country`: Country name
- `year`: Year
- `continent`: Africa, Americas, Asia, Europe, or Oceania
- `region`: World Bank region (East Asia & Pacific, Sub-Saharan Africa, etc.)
- `income_group`: World Bank income group (Low income, Lower middle income, Upper middle income, High income)
- `iso3c`: Three-letter country code
- `happiness`: Average life evaluation score, on a scale of 0–10 (higher = happier)
- `life_expectancy`: Average life expectancy at birth, in years
- `gdp_per_cap`: GDP per capita (constant 2015 US dollars)
- `gdp_per_cap_1000`: GDP per capita, in thousands of dollars
- `internet_use_prop`: Percent of the population that uses the internet (0–100)
- `urban_prop`: Percent of the population that lives in cities (0–100)
- `women_parliament_prop`: Percent of seats in the national parliament held by women (0–100)
- `unemployment_rate`: Percent of the labor force that is unemployed (0–100)
- `population_millions`: Population, in millions

```{r}
world_happiness
```


## I do: Happiness, health, and region

*I'll walk through this one on the screen—just follow along and read; you don't need to type anything*

Is national happiness explained by health, urbanization, *and* region?

### Look at the correlations

Before fitting the model, check how the outcome is related to each continuous explanatory variable, *and* how the explanatory variables are related to each other. `get_correlation()` only works with two variables at a time, but if you `select()` a bunch of numeric columns and pipe them into `cor()`, you get the correlation between every pair of them all at once (this is called a *correlation matrix*):

```{r}
world_happiness |>
  select(happiness, life_expectancy, urban_prop) |>
  cor()
```

Read it like a multiplication table: find a row and a column and the number where they cross is their correlation. The diagonal is all 1s because every variable is perfectly correlated with itself, and the top and bottom halves are mirror images of each other.

Happiness is strongly and positively correlated with both life expectancy (0.73) and urbanization (0.62). But life expectancy and urbanization are *also* strongly correlated with each other (0.68)—healthier countries tend to be more urban—so some of the explanation they each give for happiness is shared. That's why the coefficients in the multiple regression model will be different from what we'd get if we looked at each one on its own.

(`region` is categorical, so it can't go in `cor()`; correlations only work with numbers.)

### Fit the model

The model explains `happiness` with two continuous variables (sliders: `life_expectancy` and `urban_prop`) and one categorical variable (switches: `region`):

```{r}
model_region <- lm(
  happiness ~ life_expectancy + urban_prop + region,
  data = world_happiness
)

get_regression_table(model_region)
```

### Find the base case

`region` has 7 categories, but there are only 6 region coefficients in the table. The missing one is the **base case**. By default, R chooses the first region alphabetically as the base case, or East Asia & Pacific here. Every region coefficient is measured relative to East Asia & Pacific.

### Write the equation

Here's the model as an equation, with the numbers from the table:

$$
\begin{aligned}
\widehat{\text{happiness}} = &\ 0.19 + 0.062 \times \text{life expectancy} + 0.014 \times \text{urban} \\
&+ 0.30 \times \text{Europe/Central Asia} + 0.40 \times \text{Latin America/Caribbean} \\
&+ (-0.81) \times \text{Middle East/North Africa} + 0.46 \times \text{North America} \\
&+ (-0.94) \times \text{South Asia} + (-0.43) \times \text{Sub-Saharan Africa}
\end{aligned}
$$

### Interpret the coefficients

Intercept:

> In an East Asian or Pacific country with a life expectancy of 0 years and 0% of the population living in cities, average happiness would be 0.19.

This is what happens when all the sliders and switches are set to 0. In this case, no country has a life expectancy of 0 (the lowest in the data is around 55 years), so the intercept is just a mathematical starting point for the line. It doesn't really matter here.

Life expectancy (a slider):

> Taking urbanization and region into account, a one-year increase in life expectancy is associated with a 0.062 point increase in happiness, on average.

Urbanization (a slider):

> Controlling for life expectancy and region, a 1 percentage point increase in the share of people living in cities is associated with a 0.014 point increase in happiness, on average.

(That seems tiny, but remember the units: `urban_prop` goes from 0 to 100, so a 10 percentage point jump in urbanization would be associated with a 0.14 point increase in happiness.)

Latin America & Caribbean (a switch):

> Accounting for life expectancy and urbanization, happiness in Latin America & the Caribbean is 0.40 points higher than in East Asia & the Pacific, on average.

Sub-Saharan Africa (a switch):

> Taking life expectancy and urbanization into account, happiness in Sub-Saharan Africa is 0.43 points lower than in East Asia & the Pacific, on average.

::: {.callout-warning}
Some regions have very few countries in them! In this data, North America only has 2 and South Asia only has 4 in this data, so their coefficients are based on very little information. Be careful about reading too much into them.
:::


### Plot it

We can see the switches shifting the line up and down with `geom_parallel_slopes()`. This plot only uses life expectancy and region—it ignores `urban_prop`—so the lines won't match the model exactly, but it shows the general idea: one slope, shifted up or down for each region.

```{r}
ggplot(world_happiness, aes(x = life_expectancy, y = happiness, color = region)) +
  geom_point() +
  geom_parallel_slopes(se = FALSE)
```

## General process

You'll basically do this when interpreting any regression model:

1. **Find the base case**: Figure out which category of each categorical variable is missing from the table
3. **Interpret each coefficient with the template**:
    - Sliders: "Taking [the other variables] into account, a one unit increase in X is associated with a $b$ increase (or decrease) in Y, on average"
    - Switches: "Taking [the other variables] into account, Y is $b$ higher (or lower) in this category compared to the base case, on average"
4. **Make sure things make sense**: are the units right (years? percentage points? \$1,000s?), and does the intercept make any sense or is it an impossible situation?


## We do: Happiness, wealth, internet access, and continents

*We'll work through this one together—type along with me and fill in the blanks as we go*

Is national happiness explained by wealth, internet access, *and* continent?

Use `select()` and `cor()` to make a correlation matrix of `happiness`, `gdp_per_cap_1000`, and `internet_use_prop`. A few countries are missing `internet_use_prop`, so add `use = "complete.obs"` inside `cor()` to make it ignore the missing values (it's like `na.rm = TRUE`, but `cor()` uses a different name for it):

```{r}
# world_happiness |>
#   select(...) |>
#   cor(...)
```

Which variable is more strongly correlated with happiness? How correlated are GDP per capita and internet use with each other?

> ___ is more strongly correlated with happiness (r = ___). GDP per capita and internet use have a correlation of ___, which means ...

Fit a model with `lm()` that explains `happiness` with `gdp_per_cap_1000`, `internet_use_prop`, and `continent`, and look at the coefficients with `get_regression_table()`:

```{r}
# model_happiness <- lm(...)
```

What's the base case for `continent`?

> The base case is ___, because ...

Write out the model as an equation (fill in the numbers):

$$
\widehat{\text{happiness}} = b_0 + b_1 \times \text{GDP}_{1000} + b_2 \times \text{internet} + b_3 \times \text{Americas} + b_4 \times \text{Asia} + b_5 \times \text{Europe} + b_6 \times \text{Oceania}
$$

Interpret the intercept:

> In an African country with a GDP per capita of \$0 and 0% internet use, ...

Interpret the coefficient for GDP per capita:

> Controlling for internet use and continent, a \$1,000 increase in GDP per capita is associated with a ___ ...

Interpret the coefficient for internet use:

> Taking GDP per capita and continent into account, a 1 percentage point increase in internet use is associated with a ___ ...

Interpret the coefficient for Europe:

> Holding GDP per capita and internet use constant, happiness in European countries is ___ points ___ (higher/lower) compared to ___, on average.

Bonus: plot `happiness` and `gdp_per_cap_1000` with one line for each continent using `geom_parallel_slopes(se = FALSE)` (this plot ignores internet use, so the lines won't match the model exactly!):

```{r}
ggplot(data = world_happiness)
```


## You do together: Happiness, health, jobs, and income

*Do this in a group. You'll write the code and interpretations (you can have just one person do the typing if you want), and I'll walk around to help*

Is national happiness explained by health, jobs, and income?

Make a correlation matrix of `happiness`, `life_expectancy`, and `unemployment_rate` (hint: some countries are missing `unemployment_rate`!):

```{r}
# Correlation matrix here
```

How is happiness related to life expectancy and unemployment? Are life expectancy and unemployment related to each other?

> ...

Fit a model that explains `happiness` with `life_expectancy`, `unemployment_rate`, and `income_group`, and look at the coefficients with `get_regression_table()`:

```{r}
# model_income <- lm(...)
```

First, find the base case. What's the base case for `income_group`, and how do you know? (Hint: look at what the setup chunk at the top of this file does to `income_group` with `factor()`—is it still alphabetical?)

> ...

Interpret the coefficient for life expectancy:

> Controlling for unemployment and income group, ...

Interpret the coefficient for unemployment:

> ...

Interpret the coefficient for high income:

> ...

**Check with your group**: Which coefficient surprised you the most? Is the sign of the unemployment coefficient what you expected? Why or why not?

> ...


## You do alone: Women in parliament, wealth, and income

*Try this one on your own—no fill-in-the-blanks this time*

Is the share of women in parliament explained by wealth and income group?

Fit a model that explains `women_parliament_prop` with `gdp_per_cap_1000` and `income_group`, and look at the coefficients with `get_regression_table()`:

```{r}
# Fit your model and look at the coefficients here
```

What's the base case for `income_group`?

> ...

Interpret the coefficient for GDP per capita:

> ...

Interpret the coefficient for one of the income groups (your choice):

> ...

Does anything about these results surprise you? (Think about what GDP per capita and income group are each measuring, and how much they might overlap.)

> ...


## Playground

*If you finish early, do this!*

Build your own model! Pick an outcome and a few explanatory variables (try mixing continuous ones like `urban_prop`, `women_parliament_prop`, or `population_millions` with categorical ones like `region` or `income_group`). Fit the model, look at the coefficients, find the base case, and interpret each coefficient.

```{r}
# model_playground <- lm(...)
```

Interpretations:

> ...
