---
title: "Multiple regression I"
date: "2026-10-04"
---

```{r}
#| include: false

knitr::opts_chunk$set(
  fig.width = 5,
  fig.height = (5 * 0.618),
  fig.align = "center",
  fig.retina = 3,
  out.width = "75%",
  collapse = TRUE
)
```

## Readings

::: {.readings}

{{< include ../readings/_13-class.md >}}

:::

## Slides

<p style="line-height: 2.5"><a class="btn btn-success me-2" target="_blank" href="../slides/13-multiple-regression/index.html">{{< fa arrow-up-right-from-square >}} &ensp;Open slides in new window</a><a class="btn btn-success" target="_blank" href="../slides/13-multiple-regression/13-multiple-regression.pdf">{{< fa file-pdf >}} &ensp;Download PDF of slides</a></p>

```{=html}
<div class="ratio ratio-16x9">
<iframe class="slide-deck" src="../slides/13-multiple-regression/index.html" style="border: 1px solid #343a40;"></iframe>
</div>
```

## Materials for class today

- [Regression playground `.zip` project](/projects/regression-playground.zip)
- [Posit.cloud project](https://posit.cloud/spaces/809160/content/13061645) (only if you're using an iPad!)

## In-class examples

Here's the code from the two "Our Turn"s and "Your Turns"s from class today. The CSV file of the data is in the `.zip` file that you downloaded above, or you can grab if from here:

- {{< fa file-csv >}} [`world_happiness.csv`](/projects/regression-playground/data/world_happiness.csv)

If you're curious about how I made this data, you can see the R code I used to build, combine, and clean the data here:

- {{< fa brands r-project >}} [`make-world-happiness.R`](/data/make-world-happiness.R)

---

```{.r}
library(tidyverse)   # For ggplot, dplyr, and friends
library(moderndive)  # For get_correlation(), get_regression_table(), etc

# Load the World Happiness Report data and only keep 2024
world_happiness <- read_csv("data/world_happiness.csv") |>
  filter(year == 2024)
```

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(moderndive)

world_happiness <- read_csv(here::here(
  "projects",
  "regression-playground",
  "data",
  "world_happiness.csv"
)) |>
  filter(year == 2024)
```

### World happiness

This data comes from the [World Happiness Report](https://www.worldhappiness.report/) (happiness scores) and the [World Bank](https://data.worldbank.org/) (everything else). Each row is a country in 2024.

- `country`: Country name
- `year`: Year
- `continent`: Africa, Americas, Asia, Europe, or Oceania
- `region`: World Bank region (East Asia & Pacific, Sub-Saharan Africa, etc.)
- `income_group`: World Bank income group (Low income, Lower middle income, Upper middle income, High income)
- `iso3c`: Three-letter country code
- `happiness`: Average life evaluation score, on a scale of 0–10 (higher = happier)
- `life_expectancy`: Average life expectancy at birth, in years
- `gdp_per_cap`: GDP per capita (constant 2015 US dollars)
- `internet_use_prop`: Percent of the population that uses the internet (0–100)
- `urban_prop`: Percent of the population that lives in cities (0–100)
- `women_parliament_prop`: Percent of seats in the national parliament held by women (0–100)
- `unemployment_rate`: Percent of the labor force that is unemployed (0–100)
- `population_millions`: Population, in millions

```{r}
world_happiness
```


### Our turn #1: Happiness and life expectancy

Does a country's life expectancy predict its happiness?

Make a scatterplot with `life_expectancy` on the x-axis and `happiness` on the y-axis:

```{r}
#| message: false

ggplot(world_happiness, aes(x = life_expectancy, y = happiness)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Find the correlation between the two with `get_correlation()`:

```{r}
get_correlation(world_happiness, happiness ~ life_expectancy)
```

Fit a model with `lm()` and look at the coefficients with `get_regression_table()`:

```{r}
model_life_exp <- lm(
  happiness ~ life_expectancy,
  data = world_happiness
)

get_regression_table(model_life_exp)
```

Interpret the slope:

> On average, a one-year increase in life expectancy is associated with a 0.116 increase in happiness points.

Interpret the intercept:

> In a country where life expectancy is 0 years (which doesn't exist!), average happiness is -2.99.


### Your turn #1: Happiness and internet access

Does access to the internet predict national happiness?

Make a scatterplot with `internet_use_prop` on the x-axis and `happiness` on the y-axis:

```{r}
#| message: false
#| warning: false

ggplot(data = world_happiness, aes(x = internet_use_prop, y = happiness)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

Find the correlation between the two with `get_correlation()`:

```{r}
get_correlation(world_happiness, happiness ~ internet_use_prop, na.rm = TRUE)
```

Fit a model with `lm()` and look at the coefficients with `get_regression_table()`:

```{r}
model_internet <- lm(
  happiness ~ internet_use_prop, 
  data = world_happiness
)

get_regression_table(model_internet)
```

Interpret the slope:

> On average, a one percentage point increase in internet use is associated with a 0.035 point increase in happiness.

Interpret the intercept:

> In a country where the proportion of people with access to the internet is 0% (which is potentially plausible!), average happiness is 3.02.


### Our turn #2: Happiness and continents

Is happiness different across continents?

Make a plot that shows `happiness` across each `continent` (try `geom_point()` with `position = position_jitter(width = 0.15)`, or `geom_boxplot()`):

```{r}
#| include: false

set.seed(1234)
```

```{r}
ggplot(data = world_happiness, aes(x = continent, y = happiness)) + 
  geom_point(position = position_jitter(width = 0.15))
```

Find the average happiness in each continent with `group_by()` and `summarise()`:

```{r}
world_happiness |> 
  group_by(continent) |> 
  summarise(avg_happiness = mean(happiness))
```

Fit a model with `lm()` and look at the coefficients with `get_regression_table()`:

```{r}
model_happiness_continent <- lm(
  happiness ~ continent,
  data = world_happiness
)

get_regression_table(model_happiness_continent)
```

What's the base case?

> The base case, or omitted continent, is Africa.

Interpret the intercept:

> Average happiness in Africa is 4.42.

Interpret the coefficient for Europe:

> On average, happiness in Europe is 2.06 points higher than in Africa.


### Your turn #2: Internet access and continents

Is internet access different across continents?

Make a plot that shows `internet_use_prop` across each `continent`:

```{r}
#| warning: false

ggplot(data = world_happiness, aes(x = continent, y = internet_use_prop)) + 
  geom_point(position = position_jitter(width = 0.15))
```

Find the average internet use in each continent with `group_by()` and `summarise()`:

```{r}
world_happiness |> 
  group_by(continent) |> 
  summarise(avg_internet = mean(internet_use_prop, na.rm = TRUE))
```

Fit a model with `lm()` and look at the coefficients with `get_regression_table()`:

```{r}
model_internet_continent <- lm(
  internet_use_prop ~ continent,
  data = world_happiness
)

get_regression_table(model_internet_continent)
```

Interpret the intercept:

> The average proportion of internet use in Africa is 44.3%.

Interpret the coefficient for Asia:

> On average, the proportion of internet use in Asia is 36.8 percentage points higher than in Africa.
